---
title: "Introduction to {evsim}"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 9)
library(evsim)
library(dplyr)
```

This package provides the tools to simulate new EV charging session, on the basis of the **Gaussian Mixture Models (GMM)** of different EV user profiles found and modeled with [`{evprof}`](https://mcanigueral.github.io/evprof/). The EV model obtained with [`evprof::get_ev_model`](https://mcanigueral.github.io/evprof/reference/get_ev_model.html) function is an object of class `evmodel`, which contains multiple time-cycle models (e.g. Weekdays and Weekends). At the same time, each time-cycle model consists in a combination of multiple EV user profiles and each user profile consists in multiple GMM. For more information about the EV models read [this article](https://mcanigueral.github.io/evprof/articles/evmodel.html) from website of `{evprof}` package.

## Required data

The only function of `{evsim}` package necessary to simulate new EV sessions is function `simulate_sessions`, which returns a tibble of EV charging sessions. The required input arguments to run this function are:

- The EV model (object of class `evmodel`)
- Number of daily sessions for each time-cycle model
- User profiles distribution (ratio of every user profile) and specific charging power (if required)
- Charging power distribution (ratio of every different charging rate)
- Dates to simulate


### EV model

Package `{evsim}` provides an EV model of example, being the model created from California ACN data in the [California article](https://mcanigueral.github.io/evprof/articles/california.html) from `{evprof}` documentation. This model can be accessed loading the `{evsim}` package or with the following function:

```{r}
ev_model <- evsim::california_ev_model
```

Let's take a look to the information that the `print` function shows for an `evmodel` object:

```{r models obj}
print(ev_model)
```

We have 2 different time-cycles in this case (Workdays and Weekends), and each one has its corresponding user profiles. At the same time, each user profile of each time-cycle is modeled by two different models:

- Connection Models: combination of multiple Gaussian distributions of two variables (Connection Start Time and Connection Duration)
- Energy Models: combination of multiple Gaussian distributions of one variable (Energy)

Moreover, each User profile has a corresponding ratio over the total number of daily sessions. In this case, for the **Workday** time-cycle the half of charging sessions are **Vist** users and the other half **Worktime** users, while for the **Weekend** time-cycle all sessions are **Visit** users.

```{r workday models}
workday_models <- ev_model$models$user_profiles[[1]]
workday_models
```
```{r weekend models}
weekend_models <- ev_model$models$user_profiles[[2]]
weekend_models
```



### Number of sessions per day

To set the number of sessions per day to function `simulate_sessions()` we need to pass a tibble with variables `time_cycle` (names corresponding to `evmodel$models$time_cycle`) and `n_sessions` (number of daily sessions per day for each time-cycle model).
For example:

```{r n_sessions}
sessions_day <- tibble(
  time_cycle = ev_model$models$time_cycle,
  n_sessions = c(150, 50)
)
sessions_day
```


### User profiles distribution

In the "EV model" section we have seen that inside the `ev_model` object there is already a default distribution of user profiles, each one with the corresponding ratio. These ratio values were obtained from the clustering process done by `{evprof}` package and represent the presence of every user profile over the total number of sessions clustered.

However, we may want to simulate different scenarios or use cases where these ratios differ from the current ones. For example, to simulate the EV demand in an industrial area it could be interesting to prioritize the presence of Worktime sessions over Visit sessions. This can be done by setting a new user profiles' distribution tibble, which must contain the variables `time_cycle`, `profile` and `ratio`.

The default values for these variables, included within the model, can be obtained with the function `get_user_profiles_distribution`. It is not mandatory that all time-cycles and user profiles appear in the `user_profiles` tibble. The rule is simple: only time cycles and user profiles appearing in the `user_profiles` tibble will be simulated.

Moreover, we can specify the **charging power** of a concrete user profile. This is useful in some scenarios where there is a type of vehicle (e.g. truck) that charges at different rate than the rest of vehicles. This can be done with the column `power` in the `user_profiles` tibble, setting the power values in kW. All user profiles without a specific `power` value will be simulated with a random charging power obtained according to the charging power distribution (see next section).

For example, let's say we want to simulate a  **Workday** distribution with 80% of sessions being Worktime and 20% of sessions being Visit, and considering that all Worktime sessions charge with three-phase AC vehicles with 16A connection (11 kW). Then our `user_profiles` tibble should be the following one:

```{r}
user_profiles <- tibble(
  time_cycle = c('Workday', 'Workday', 'Weekend'),
  profile = c('Visit', 'Worktime', 'Visit'),
  ratio = c(0.2, 0.8, 1),
  power = c(NA, 11, NA)
)

user_profiles
```


### Charging powers distribution

The charging power is not a variable modeled by Gaussian Mixture Models (GMM) like the connection variables or the energy. However, the energy GMM can be built by charging power separately (see [User profiles modelling](https://mcanigueral.github.io/evprof/articles/evprof.html#user-profiles-modeling) on `{evprof}` website), since in general the energy charged depends on the charging power (the higher the charging power the higher the energy demand).

The purpose of letting the charging power out of the GMM is that it is a variable that depends on the EV model and the charging point, so it may change between scenarios. For example, we may want to simulate a scenario where all vehicles have a three-phase connection (11 kW) instead of the current scenario with still a lot of vehicles charging with a single phase (3.7 kW). Therefore, it was convenient to set an option to simulate a specific distribution of charging powers for the simulated EV fleet. This is done with a `charging_powers` tibble with variables `power` and `ratio`. In the public charging infrastructure the most common and standard power rates are:

* 3.7 kW AC (single-phase 16A)
* 7.3 kW AC (two-phase 16A)
* 11 kW AC (three-phase 16A)

If the energy GMM were built according to the charging power, it is recommended that the `power` values in the `charging_powers` tibble match the original charging power values. These can be found inside the models. For the example `ev_model` at issue, we see that the Workday energy models of the `Visit` user profile are not built by charging power since the `charging_rate` column shows an `Unknown` value:

```{r}
workday_models$energy_models[[1]] # Visit profile energy models
```

In this case, the same energy GMM will be used for all `power` values. However, if we had energy models for 3.7 kW and 11 kW separately, and we want to simulate 22 kW sessions, the models used for the 22 kW sessions will be the 11 kW models since they are the best approximate solution. 

To follow the example, we can set the following charging power distribution:
```{r charging rates}
charging_powers <- tibble(
  power = c(3.7, 7.3, 11),
  ratio = c(0.2, 0.4, 0.4)
)
charging_powers
```

This complements the variable `power` from the `user_profiles` tibble (see previous section). This `charging_powers` tibble will be used to assign a charging power to all user profiles' sessions without a `power` specified in `user_profiles`.


  
### Datetime values

The function `simulate_sessions()` automatically selects the time-cycle that fits to a specific date. This is why we have to pass a vector of the dates that we want to simulate. For example:

```{r dates}
dates_sim <- seq.Date(from = as.Date('2019-09-10'), to = as.Date('2019-09-15'), by = '1 day')
dates_sim
```

Another datetime argument that the simulation function needs is the time resolution of the datetime variables of the sessions (connection start, connection end, charging end, ...). For this example we will set a time resolution of 15 minutes, so the argument will be `resolution = 15`.


## Simulation

We have now all required data and we are ready to simulate the new sessions:

```{r estimate sessions}
sessions_estimated <- simulate_sessions(
  ev_model,
  sessions_day, 
  user_profiles,
  charging_powers,
  dates_sim,
  resolution = 15
)

head(sessions_estimated)
```

We have obtained a total of 700 sessions for our 7 simulation dates (i.e. 100 sessions/day), each session corresponding to a specific charging profile. We can notice that the datetime variables have a resolution of 15 minutes according to our configuration, and the energy required corresponds to product of `Power` and `ChargingHours`. 

We can check that the number of sessions of each power rate corresponds to our configuration:

```{r}
sessions_estimated %>% 
  group_by(Profile, Power) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n)*100)
```



## EV demand

Finally, we can calculate the estimated demand with function `get_demand()`. This function accepts an optional argument, `dttm_seq`, which will be the datetime sequence of the time-series demand data frame returned by the function. We will set the datetime sequence from our first to last date with a **resolution of 15 minutes** and the timezone of the original model (using the `lubridate::with_tz()` function):

```{r dttm_seq}
dttm_seq <- seq.POSIXt(
  from = as.POSIXct('2019-09-10'), 
  to = as.POSIXct('2019-09-16'), 
  by = '15 min'
) %>% 
  lubridate::with_tz(
    ev_model$metadata$tzone
  )
```

```{r simualted demand}
estimated_demand <- sessions_estimated %>% 
  get_demand(dttm_seq)
```

We can also visualize the estimated demand in an HTML plot using the function `plot_ts`:

```{r plot, echo=TRUE, fig.width=9}
estimated_demand %>% 
  plot_ts(ylab = 'Power demand (kW)', fillGraph = T, stackedGraph = T)
```


