---
title: "Multi-core processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, error = F)
devtools::load_all() # load evsim
library(dplyr)
library(lubridate)
```

The time-series calculations done by functions `get_demand` and `get_n_connections` can be done by multiple time windows in parallel. At the same time, in computer science, a common practice to reduce computation times is parallel processing when the computer has multiple CPUs (i.e. multi-core processing). In parallel processing, a task is divided into multiple sub-tasks and assigned to multiple cores to make de calculations in parallel. For parallel processing in R we make use of the R base package `parallel`.

We can discover the number of physical CPUs in our computer with function `parallel::detectCores`:
```{r}
n_cores <- parallel::detectCores(logical = FALSE)
print(n_cores)
```

We set the parameter `logical = FALSE` because it is important to know the **physical** CPUs, not the logical (or virtual/thread) ones seen by the OS. More information about this parameter can be found in `parallel::detectCores` documentation.

Below, an example of calculating the demand of 10.000 EV charging sessions with different number of cores:

```{r sessions, echo = T, eval = F}
cores_time <- tibble(
  cores = 1:n_cores,
  time = 0
)
for (mcc in cores_time$cores) {
  results <- system.time(
    evsim::california_ev_sessions %>%
       sample_n(10000) %>%
       mutate(Profile = "All") %>%
       get_demand(by = "Profile", resolution = 60, mc.cores = mcc)
  )
  cores_time$time[mcc] <- as.numeric(results[3])
}
```

```{r saved data, echo=F, eval = T}
# saveRDS(cores_time, "multi_processing_data/cores_time.RDS")
cores_time <- readRDS("multi_processing_data/cores_time.RDS")
```


And below there's the plot of the results, were we can see that, for this computer, the optimal number of cores to use is 4:
```{r}
library(ggplot2)
cores_time %>%
  ggplot(aes(x = cores, y = time)) +
  geom_col() + 
  labs(x = "Number of cores", y = "Computation time (seconds)")
```





```{r}
evmodel <- evsim::california_ev_model
sessions_day 
ev_models <- prepare_model(evmodel[["models"]], sessions_day, user_profiles)
connection_log <- evmodel[['metadata']][['connection_log']]
energy_log <- evmodel[['metadata']][['energy_log']]
tzone_model <- evmodel[['metadata']][['tzone']]

dates_dttm <- round_date(as_datetime(dates, tz = tzone_model), unit = 'day')

simulated_sessions <- map_dfr(
  dates_dttm,
  ~ get_day_sessions(.x, ev_models, connection_log, energy_log, charging_powers)
)
```





